---
output:
  github_document:
    html_preview: false
---

<!-- README.md and index.md are generated from README.Rmd. Please edit that file. -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

set.seed(20230702)

clean_output <- function(x, options) {
  # Side effect
  usethis::use_build_ignore("index.md")

  x <- gsub("0x[0-9a-f]+", "0xdeadbeef", x)
  x <- gsub("dataframe_[0-9]*_[0-9]*", "      dataframe_42_42      ", x)
  x <- gsub("[0-9]*\\.___row_number ASC", "42.___row_number ASC", x)

  index <- x
  index <- gsub("─", "-", index)
  index <- strsplit(paste(index, collapse = "\n"), "\n---\n")[[1]][[2]]
  writeLines(index, "index.md")

  x <- gsub('(`vignette[(]"([^"]+)"[)]`)', "[\\1](https://rprojroot.r-lib.org/articles/\\2.html)", x)
  x <- fansi::strip_sgr(x)
  x
}

options(
  cli.num_colors = 256,
  cli.width = 71,
  width = 71,
  pillar.bold = TRUE,
  pillar.max_title_chars = 5,
  pillar.min_title_chars = 5,
  pillar.max_footer_lines = 12,
  conflicts.policy = list(warn = FALSE)
)

local({
  hook_source <- knitr::knit_hooks$get("document")
  knitr::knit_hooks$set(document = clean_output)
})

rlang::local_interactive(FALSE)
```

# [rprojroot](https://rprojroot.r-lib.org/)

<!-- badges: start -->
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html)
[![rcc](https://github.com/r-lib/rprojroot/workflows/rcc/badge.svg)](https://github.com/r-lib/rprojroot/actions)
[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/rprojroot)](https://cran.r-project.org/package=rprojroot)
[![Codecov test coverage](https://codecov.io/gh/r-lib/rprojroot/branch/main/graph/badge.svg)](https://app.codecov.io/gh/r-lib/rprojroot?branch=main)
<!-- badges: end -->

This package helps accessing files relative to a *project root* to [stop the working directory insanity](https://gist.github.com/jennybc/362f52446fe1ebc4c49f).
It is a low-level helper package for the [here](https://here.r-lib.org/) package.

```{r}
library(rprojroot)
```


## Example

The rprojroot package works best when you have a "project": all related files contained in a subdirectory that can be categorized using a strict criterion.
Let's create a package for demonstration.

```{r}
dir <- tempfile()
pkg <- usethis::create_package(dir)
```

R packages satisfy the `is_r_package` criterion.
A criterion is an object that contains a `find_file()` function.
With `pkg` as working directory, the function works like `file.path()`, rooted at the working directory:

```{r}
setwd(pkg)
is_r_package
is_r_package$find_file()
is_r_package$find_file("tests", "testthat")
```

This works identically when starting from a subdirectory:

```{r}
setwd(file.path(pkg, "R"))
is_r_package$find_file()
is_r_package$find_file("tests", "testthat")
```

There is one exception: if the first component passed to `find_file()` is already an absolute path.
This allows safely applying this function to paths that may be absolute or relative:

```{r}
setwd(file.path(pkg, "R"))
path <- is_r_package$find_file()
is_r_package$find_file(path, "tests", "testthat")
```


As long as you are sure that your working directory is somewhere inside your project, you can retrieve the project root.


## Installation and further reading

Install the package from CRAN:

``` r
install.package("rprojroot")
```

See the [documentation](https://rprojroot.r-lib.org/articles/rprojroot.html) for more detail.

---

## Code of Conduct

Please note that the rprojroot project is released with a [Contributor Code of Conduct](https://rprojroot.r-lib.org/CODE_OF_CONDUCT.html).
By contributing to this project, you agree to abide by its terms.
