---
output: cynkratemplate::readme_document
---

<!-- README.md and index.md are generated from README.Rmd. Please edit that file. -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

set.seed(20230702)

clean_output <- function(x, options) {

  # The demo package below is created inside the session's temporary
  # directory, whose name is random and so different on every render.
  # usethis echoes that directory in more than one spelling: the "Creating"
  # line repeats the path as it was passed in, while "Setting active project"
  # reports the resolved one.
  # On macOS the two differ, because `/var` is a symlink to `/private/var`,
  # and on Windows a backslashed spelling shows up as well.
  # Substitute every spelling, longest first, so that a short one is never
  # replaced inside a longer one and leaves a stray prefix behind.
  temp_dirs <- unique(c(
    tempdir(),
    chartr("\\", "/", tempdir()),
    normalizePath(tempdir(), winslash = "/"),
    normalizePath(tempdir(), winslash = "\\")
  ))
  for (temp_dir in temp_dirs[order(nchar(temp_dirs), decreasing = TRUE)]) {
    x <- gsub(temp_dir, "/tmp/RtmpXXXXXX", x, fixed = TRUE)
  }
  # The demo package's DESCRIPTION records whichever roxygen2 wrote it:
  # `RoxygenNote` on a release build, `Config/roxygen2/version` on a
  # development one. That makes the rendered README differ by machine rather
  # than by run, so CI would rewrite it on every re-render. Dropped outright --
  # the demo package's roxygen version is not what this README demonstrates.
  # Matched loosely because this hook runs before the colour is stripped, so
  # the field name may be wrapped in SGR escapes.
  x <- gsub("(?m)^#>[^\n]*(RoxygenNote|Config/roxygen2/version)[^\n]*\n", "", x, perl = TRUE)

  x <- gsub("0x[0-9a-f]+", "0xdeadbeef", x)
  x <- gsub("dataframe_[0-9]*_[0-9]*", "      dataframe_42_42      ", x)
  x <- gsub("[0-9]*\\.___row_number ASC", "42.___row_number ASC", x)

  x
}

options(
  cli.num_colors = 256,
  # Pinned so that a render under a non-UTF-8 locale keeps cli's Unicode check
  # marks, instead of downgrading each of them to `v` and churning a dozen
  # lines of output.
  cli.unicode = TRUE,
  cli.width = 71,
  width = 71,
  pillar.bold = TRUE,
  pillar.max_title_chars = 5,
  pillar.min_title_chars = 5,
  pillar.max_footer_lines = 12,
  conflicts.policy = list(warn = FALSE)
)

# `usethis::create_package()` fills the demo package's DESCRIPTION from this
# option, so whoever renders the README decides what the front page shows.
# Pinned here to the data it is meant to show, rather than left to the
# maintainer's `.Rprofile`, which also contributed a `Date:` of the render day
# and this package's own `URL:` to a DESCRIPTION that is not this package's.
# The umlaut is written as an escape, so that the field reads the same however
# the rendering session interprets the source file.
options(
  usethis.description = list(
    `Authors@R` = 'person("Kirill", "M\u00fcller", email = "kirill@cynkra.com", role = c("aut", "cre"), comment = c(ORCID = "0000-0002-1416-3412"))',
    License = "MIT + file LICENSE"
  )
)

local({
  hook_source <- knitr::knit_hooks$get("document")
  knitr::knit_hooks$set(document = clean_output)
})

rlang::local_interactive(FALSE)
```

# [rprojroot](https://rprojroot.r-lib.org/)

<!-- badges: start -->
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html)
[![rcc](https://github.com/r-lib/rprojroot/workflows/rcc/badge.svg)](https://github.com/r-lib/rprojroot/actions)
[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/rprojroot)](https://cran.r-project.org/package=rprojroot)
[![Codecov test coverage](https://codecov.io/gh/r-lib/rprojroot/graph/badge.svg)](https://app.codecov.io/gh/r-lib/rprojroot/tree/main)
<!-- badges: end -->

This package helps accessing files relative to a *project root* to [stop the working directory insanity](https://gist.github.com/jennybc/362f52446fe1ebc4c49f).
It is a low-level helper package for the [here](https://here.r-lib.org/) package.

```{r}
library(rprojroot)
```


## Goals and non-goals

rprojroot aims to:

- Find the root of a project by walking up from a starting directory until a directory matches a criterion.
- Ship criteria for the project layouts people already have,
  from `is_r_package` and `is_rstudio_project` to `is_git_root` and `is_renv_project`;
  see `criteria` for the full list.
- Let you state your own criterion with `root_criterion()` or `has_file()`,
  and combine criteria with `|`.
- Build paths below the root that stay correct wherever the working directory happens to be,
  with `find_file()` and `make_fix_file()`.
- Install with no dependencies beyond base R.

It is explicitly not trying to:

- Be the package you reach for in an interactive session:
  the [here](https://here.r-lib.org/) package sits on top of rprojroot for that, and is intended for interactive use only.
- Guess which criterion applies to your project:
  `find_root()` takes the criterion as its first argument, and raises an error when no directory above the starting path matches it.
- Determine the path of the currently running script:
  `thisfile()` is soft-deprecated, and these functions are now available in the whereami package.

## Example

The rprojroot package works best when you have a "project":
all related files contained in a subdirectory that can be categorized using a strict criterion.
Let's create a package for demonstration.

```{r}
# A fixed name, so the demo package is not called `file1a2b3c` in the output.
dir <- file.path(tempdir(), "demopkg")
pkg <- usethis::create_package(dir)
```

R packages satisfy the `is_r_package` criterion.
A criterion is an object that contains a `find_file()` function.
With `pkg` as working directory, the function works like `file.path()`, rooted at the working directory:

```{r}
setwd(pkg)
is_r_package
is_r_package$find_file()
is_r_package$find_file("tests", "testthat")
```

This works identically when starting from a subdirectory:

```{r}
setwd(file.path(pkg, "R"))
is_r_package$find_file()
is_r_package$find_file("tests", "testthat")
```

There is one exception: if the first component passed to `find_file()` is already an absolute path.
This allows safely applying this function to paths that may be absolute or relative:

```{r}
setwd(file.path(pkg, "R"))
path <- is_r_package$find_file()
is_r_package$find_file(path, "tests", "testthat")
```


As long as you are sure that your working directory is somewhere inside your project, you can retrieve the project root.


## Installation and further reading

Install the package from CRAN:

``` r
install.package("rprojroot")
```

Or the development version from GitHub with:

``` r
# install.packages("pak")
pak::pak("r-lib/rprojroot")
```

See the [documentation](https://rprojroot.r-lib.org/articles/rprojroot.html) for more detail.

---

## Code of Conduct

Please note that the rprojroot project is released with a [Contributor Code of Conduct](https://rprojroot.r-lib.org/CODE_OF_CONDUCT.html).
By contributing to this project, you agree to abide by its terms.
