---
output: downlit::readme_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)

rlang::local_interactive(FALSE)
```

# [rprojroot](https://r-lib.github.io/rprojroot)

<!-- badges: start -->
[![rcc](https://github.com/r-lib/rprojroot/workflows/rcc/badge.svg)](https://github.com/r-lib/rprojroot/actions)
[![codecov.io](https://codecov.io/github/r-lib/rprojroot/coverage.svg?branch=master)](https://codecov.io/github/r-lib/rprojroot?branch=master)
[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/rprojroot)](https://cran.r-project.org/package=rprojroot)
<!-- badges: end -->

This package helps accessing files relative to a *project root* to [stop the working directory insanity](https://gist.github.com/jennybc/362f52446fe1ebc4c49f).
It is a low-level helper package for the [here](https://here.r-lib.org/) package.

```{r}
library(rprojroot)
```


## Example

The rprojroot package works best when you have a "project": all related files contained in a subdirectory that can be categorized using a strict criterion.
Let's create a package for demonstration.

```{r}
dir <- tempfile()
pkg <- usethis::create_package(dir)
usethis::proj_set(pkg)
```

R packages satisfy the `is_r_package` criterion.
A criterion is an object that contains a `find_file()` function:

```{r}
is_r_package
is_r_package$find_file(pkg)
is_r_package$find_file(pkg, "tests", "testthat")
```

This also works when starting from a subdirectory:

```{r}
is_r_package$find_file(file.path(pkg, "tests", "testthat"))
```

As long as you are sure that you have a path somewhere inside a project, you can retrieve the project root.


## Installation and further reading

Install the package from CRAN:

``` r
install.package("rprojroot")
```

See the [documentation](https://r-lib.github.io/rprojroot/articles/rprojroot.html) for more detail.

---

## Code of Conduct

Please note that the rprojroot project is released with a [Contributor Code of Conduct](https://rprojroot.r-lib.org/CODE_OF_CONDUCT.html).
By contributing to this project, you agree to abide by its terms.
