---
output:
  md_document:
    variant: markdown_github
knit: (function(inputFile, encoding) {
      rmarkdown::render(
        inputFile, encoding=encoding,
        output_file=rprojroot::find_root_file("README.md", criterion = "DESCRIPTION")) })
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

[rprojroot](https://krlmlr.github.io/rprojroot) [![Travis-CI Build Status](https://travis-ci.org/krlmlr/rprojroot.svg?branch=master)](https://travis-ci.org/krlmlr/rprojroot) [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/krlmlr/rprojroot?branch=master&svg=true)](https://ci.appveyor.com/project/krlmlr/rprojroot) [![codecov.io](https://codecov.io/github/krlmlr/rprojroot/coverage.svg?branch=master)](https://codecov.io/github/krlmlr/rprojroot?branch=master) [![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/rprojroot)](https://cran.r-project.org/package=rprojroot)
=================================================================================================================================================================================================================================================================================================================================================

This package helps accessing files relative to a *project root* to [stop the working directory insanity](https://gist.github.com/jennybc/362f52446fe1ebc4c49f).

## Example

The source for this text is in the [`readme` subdirectory](https://github.com/krlmlr/rprojroot/tree/master/readme):

```{r}
basename(getwd())
```

How do we access the package root? In a robust fashion? Easily:

```{r}
dir(rprojroot::find_root("DESCRIPTION"))
```


## Installation and further reading

Install from GitHub:

``` r
devtools::install_github("krlmlr/rprojroot")
```

See the [vignette](http://krlmlr.github.io/rprojroot/vignettes/rprojroot.html) for more detail.
